Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
suppress negative whitespace in \Delta\omega
\boldsymbol{\psi} Fig 1

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006,sainath2015interspeech}.
In addition, Gammatone wavelets have proven to outperform Morlet wavelets on a benchmark of supervised musical instrument classification from scattering coefficients \cite{lostanlen2017phd}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

@@

\[
\dfrac{\vert\boldsymbol{\hat{\psi}}(\omega)\vert}{\vert\boldsymbol{\hat{\psi}}(\xi)\vert}=\dfrac{\omega}{\xi}\times\left(1+\dfrac{\left(\omega-\xi\right)^{2}}{\sigma^{2}}\right)^{-N/2}=r,
\]
where the magnitude cutoff $r$ is most
often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\!\omega=\omega-\xi$.
Raising the above equation to the power $N/2$ yields the following:
\[
\left(1+\dfrac{\Delta\!\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\!\omega^{2}}{\alpha^{2}}\right).
\]
Since $\Delta\!\omega\ll\xi$, we may approximate the left-hand side
with a first-order Taylor expansion. This leads to a quadratic equation
of the variable $\Delta\negmedspace\omega$:
\[
\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\!\omega^{2}-\dfrac{2}{N\xi}\times\Delta\!\omega+\left(r^{2/N}-1\right)=0.
\]
The discriminant of the above equation is:
\[
D=4\times\left(\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}\right),
\]
which is a positive number as long as $r<1$. The bandwidth $B$ of
$\boldsymbol{\hat{\psi}}$ is given by the difference between the
two solutions of the quadratic equation, that is:
\[
B=\dfrac{2\sigma^{2}}{r^{2/N}}\times\sqrt{\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}}.
\]
Now, let us express the parameter $\alpha$ as a function of some
required bandwidth $B$ at some cutoff threshold $r$. After having
raised the above to its square and rearranged the terms, we obtain
another quadratic equation, yet of the variable $\mbox{\ensuremath{\alpha}}^{2}$:
\[
\dfrac{4}{r^{4/N}N^{2}\xi^{2}}\sigma^{4}+\dfrac{4\times\left(1-r^{2/N}\right)}{r^{2/N}}\sigma^{2}-B^{2}=0
\]
We multiply the equation by $\dfrac{r^{2/N}}{4\times\left(1-r^{2/N}\right)}\neq0:$
\[
\dfrac{1}{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}\sigma^{4}+\sigma^{2}-\dfrac{r^{2/N}B^{2}}{4\times\left(1-r^{2/N}\right)}=0
\]
This leads to defining $\sigma^{2}$ as the unique positive root of
the above polynomial:
\[
\sigma^{2}=\dfrac{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}{2}\times\left(\sqrt{1+\dfrac{B^{2}}{\left(1-r^{2/N}\right)^{2}N^{2}\xi^{2}}}-1\right).
\]
If the filter bank has to be approximately orthogonal, we typically
set $B$ to $B=\left(1-2^{-1/Q}\right)\times\xi$.

>>>>

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\boldsymbol{\psi}(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006,sainath2015interspeech}.
In addition, Gammatone wavelets have proven to outperform Morlet wavelets on a benchmark of supervised musical instrument classification from scattering coefficients \cite{lostanlen2017phd}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

@@

\[
\dfrac{\vert\boldsymbol{\hat{\psi}}(\omega)\vert}{\vert\boldsymbol{\hat{\psi}}(\xi)\vert}=\dfrac{\omega}{\xi}\times\left(1+\dfrac{\left(\omega-\xi\right)^{2}}{\sigma^{2}}\right)^{-N/2}=r,
\]
where the magnitude cutoff $r$ is most
often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\omega=\omega-\xi$.
Raising the above equation to the power $N/2$ yields the following:
\[
\left(1+\dfrac{\Delta\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\omega^{2}}{\alpha^{2}}\right).
\]
Since $\Delta\omega\ll\xi$, we may approximate the left-hand side
with a first-order Taylor expansion. This leads to a quadratic equation
of the variable $\Delta\negmedspace\omega$:
\[
\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\omega^{2}-\dfrac{2}{N\xi}\times\Delta\omega+\left(r^{2/N}-1\right)=0.
\]
The discriminant of the above equation is:
\[
D=4\times\left(\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}\right),
\]
which is a positive number as long as $r<1$. The bandwidth $B$ of
$\boldsymbol{\hat{\psi}}$ is given by the difference between the
two solutions of the quadratic equation, that is:
\[
B=\dfrac{2\sigma^{2}}{r^{2/N}}\times\sqrt{\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}}.
\]
Now, let us express the parameter $\alpha$ as a function of some
required bandwidth $B$ at some cutoff threshold $r$. After having
raised the above to its square and rearranged the terms, we obtain
another quadratic equation, yet of the variable $\mbox{\ensuremath{\alpha}}^{2}$:
\[
\dfrac{4}{r^{4/N}N^{2}\xi^{2}}\sigma^{4}+\dfrac{4\times\left(1-r^{2/N}\right)}{r^{2/N}}\sigma^{2}-B^{2}=0
\]
We multiply the equation by $\dfrac{r^{2/N}}{4\times\left(1-r^{2/N}\right)}\neq0:$
\[
\dfrac{1}{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}\sigma^{4}+\sigma^{2}-\dfrac{r^{2/N}B^{2}}{4\times\left(1-r^{2/N}\right)}=0
\]
This leads to defining $\sigma^{2}$ as the unique positive root of
the above polynomial:
\[
\sigma^{2}=\dfrac{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}{2}\times\left(\sqrt{1+\dfrac{B^{2}}{\left(1-r^{2/N}\right)^{2}N^{2}\xi^{2}}}-1\right).
\]
If the filter bank has to be approximately orthogonal, we typically
set $B$ to $B=\left(1-2^{-1/Q}\right)\times\xi$.
